### 2019-09-02 paper/learning/notes.tex
updated

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2018

% ready for submission
% \usepackage{neurips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage{iclr2020_conference,times}
% to compile a camera-ready version, add the [final] option, e.g.:
     % \usepackage[final]{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{graphicx}
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{listings}
\usepackage{tabularx}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{tikz,pgf}
\usepackage{todonotes}
\usepackage{tikz-cd}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{wrapfig}
\usepackage{amsmath,amsthm}
\usepackage{amssymb}
\usepackage{cancel}
\usepackage{wrapfig}
\usepackage{amssymb}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\tikzset{>=latex}
\usetikzlibrary{bayesnet}
\usetikzlibrary{arrows,shapes,automata}
\usetikzlibrary{positioning}
\lstnewenvironment{cpp}
{\lstset{language=C++,
  basicstyle=\footnotesize,
  keywordstyle=\color{blue}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  commentstyle=\color{green}\ttfamily,
  morecomment=[l][\color{magenta}]{\#},
  breaklines = true,
  stringstyle=\color{blue}\ttfamily, % typewriter type for strings, blue 
  showstringspaces=false,
  postbreak=\raisebox{0ex}[0ex][0ex]{\color{red}$\hookrightarrow$\space},
}}{}
\newcommand{\n}[1]{\ensuremath{n_{#1}}}
\newcommand{\rv}[1]{\ensuremath{{#1}_{1:n_{#1}}}}
\newcommand{\rvv}[1]{\ensuremath{{#1}_{1:n_{#1'}}}}
\newcommand{\model}[1]{\ensuremath{\mathcal{M}{(#1)}}}
\newcommand{\simu}[1]{\ensuremath{\Omega(\model{#1})}}
\newcommand{\micro}[2]{\ensuremath{\phi_{#1}(#2)}}
\newcommand{\pdf}[1]{\ensuremath{\mathcal{P}_{d}(#1)}}
\newcommand{\pmf}[1]{\ensuremath{\mathcal{P}_{m}(#1)}}
\newcommand{\add}[1]{\ensuremath{A_{#1}}}
\newcommand{\adds}[2]{\ensuremath{A_{#1}_{#2}}}
\newcommand{\expect}[1]{\ensuremath{\mathbb{E}[#1]}}
\newcommand{\trace}[1]{\ensuremath{\mathcal{T}_{#1}}}
\newcommand{\traces}[2]{\ensuremath{\textit{tr}^{#1}_{#2}}}
\newcommand{\stoc}[2]{\ensuremath{f^{#1}_{#2}(\rvv{x})}}
% \usepackage[binary-units=true]{siunitx}
\title{Amortization of Sub-programs in Simulators }%\\ for Fast Simulator Inference}

% neurlps classification
% Primary: Applications
% Secondary cat:Data, Challenges, Implementations, and Software
% or Hijacking Malaria Simulators with Probabilistic Programming

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.
\newcommand{\bg}[1]{~{{[{\it \textcolor{red}{{\bf BG:} #1}}]}}}
\newcommand{\cs}[1]{~{{[{\it \textcolor{red}{{\bf CS:} #1}}]}}}
\newcommand{\ag}[1]{~{{[{\it \textcolor{red}{{\bf AG:} #1}}]}}}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2019}

% \lstset{language=C++,
%                 basicstyle=\footnotesize,
%                 keywordstyle=\color{blue}\ttfamily,
%                 stringstyle=\color{red}\ttfamily,
%                 commentstyle=\color{green}\ttfamily,
%                 morecomment=[l][\color{magenta}]{\#},
%                 breaklines = true
% }
\pgfplotsset{width=7cm,compat=1.8}
\author{%
Paper ID: 1010 }
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
%}
\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}

  In this work we develop both a formalism and framework for amortizing sub-programs
  in simulators leveraging state-of-the-art probabilistic programming and variational inference techniques.
  In many real world scenarios, simulators are deployed to model computationally expensive
  and hard, or expensive, to observe scenarios.
  The simulators may represent a set of events, or a population of interacting agents. 
  Recent work by \citet{baydin2018efficient} provided a framework for converting event-based simulators 
  to probabilistic programs, without having to re-implement the simulator within a given probabilistic
  programming system (PPS).
  However, issues arise with both memory and inference when \emph{nested}, or sub-programs within the simulator are simulating 
  some random process. 
  We demonstrate that by amortizing such sub-programs not only do we significantly reduce memory consumption, but
  we also enable inference to be performed in a statistically valid framework. 
\end{abstract}

\section{Introduction}
\label{sec:intro}
Simulators arise in a myriad of scientific and industrial settings; epidemiology, physics, engineering, climate modelling and so forth [add citations].
One interesting class of simulators are population-based simulators, used in epidemiology, reinforcement learning
and financial modelling [add citations].
Performing inference in such population-based simulators is challenging, because we have nested-programs, 
which come in the form of rejection sampling blocks. 
In addition to this, each population member can traverse a number of different paths and 
can interact with other members, in certain instances, which can lead to a combinatorial number of paths 
to explore. 
Equally, in many simulated settings we cannot correctly construct the functional form of the density of the desired program,
since we do not have explicit access to the density of the \emph{program}.
Where density refers to the joint density constructed by the program, which in this case is represented by the simulator.
This inability to evaluate the joint density has spawned research into how to perform 
inference in likelihood free settings~\cite{greenberg2019automatic, mohamed2016learning, papamakarios2018sequential,brehmer2018mining}
for conditional density estimation. However, whilst many of these methods look to generate a surrogate 
for the entire simulator, we propose amortizing only computational expensive sub-components 
of the simulator, which we can automate with probabilistic programming. 
This enables three things. 
1) It reduces the loss of information, as when simulators are amortized with in current
settings we require, a potentially infinite number of samples, to provide a useful learnt representation of the simulator. 
2) It increases computational efficiency, whilst maintaining expressivity. 
3) It converts expensive rejection sampling loops into to learnt functions that can be 
easily sampled from and used to construct a functional form of the likelihood, that enables the 
target density to be fully constructed, enabling one to leverage a large class of sample efficient inference schemes, namely MCMC.

% To model a single population member quantum-based methods leveraging 
% linear superposition and entanglement provide an ability to explore multiple paths simultaneously maintaining 
% the whole program, \emph{state}, in one iteration.
%  Unfortunately, analytically writing the functional 
% form of the state is, in this instance, impossible.   
% 
In particular, we focus on combining the hijacking of simulators via probabilistic programming~\cite{baydin2018efficient} and 
state-of-the-art amortization schemes for density estimation \textbf{ Do home work on current papers}.
This enables such a process to happen in automatic way, simplifying existing set-ups, whilst providing
more expressive modelling capabilities. 
We are able to demonstrate this by replacing sub-programs, in particular, rejection sampling blocks.
Rejection sampling blocks arise in many non-standard forms within simulator to simulate densities that have no well defined cumulative density 
function, nor no known process to efficiently sample from such a density exists. 
By amortizing the rejection sampling blocks, during the inference stage when a simulator is run \emph{forword} we 
can replace such blocks with their leant surrogate, from which we can then construct the entire 
density of the simulator.
This can then be utilized, \emph{correctly}, contrary to ~\cite{baydin2018efficient},
by sample efficient Markov chain Monte Carlo inference schemes, such as Hamiltonian Monte Carlo~(HMC),
random-walk Metropolis Hastings~(RMH), Light-weight Metropolis Hastings~(LMH) and other schemes.
Whilst also significantly reducing memory consumption and computational running time of 
both the \texttt{forward()} runs of the simulator and the inferences performed via a probabilistic programming system.
This is because evaluating amortized functions 
is cheap and the reduction in rejection sampling blocks significantly reduces the number of 
unnecessary latent variables tracked by the system developed in \cite{baydin2018efficient}.

% Without the amortization procedure, as updates are made sequentially, MCMC methods are typically not-well suited this type of recursive
% estimation problem. 
% As each time we get a new data point $y_{t+1}$ we have to run new MCMC simulations 
% for $P(x_{0:t_1}| y_{0:t+1})$, we also cannot directly reuse the samples $\{x^{i}_{0:t}\}$. 
% However, as we can only run the simulator $\texttt{forward()}$ we cannot return back to a previous state and re-run. 
% So adapting existing inference techniques is the only viable option. 
% MCMC methods can be effective with a well designed amortized inference procedure (\textbf{Note: This is not guaranteed}).
% Non-MCMC based methods such as importance sampling (IS) and sequential monte carlo (SMC), may also be
% sensible inference engines, although we shall discuss the limitations within the simulator context in the proceeding sections.

However, to construct this hybrid procedure requires both substantial engineering contributions 
and a description of the mathematical framework for these non-standard probabilistic programs. 
We provide both of these in this paper. As to implement this in practice we are required to do the following:
 \begin{itemize}
  \item A mathematical framework to amortize sub-programs within simulators. 
    \item An automated training procedure for any given simulator.
    \item Extend the protocols of existing probabilistic programming systems that hijack simulators, 
    to enable only the necessary raw samples from the desired rejection sampling addresses to be collected for training.
    \item A process at inference time that replaces those given addresses with a learnt surrogate.
        % \item Finer grained control over the actual latent variables of interest, to avoid unnecessary memory consumption and ensure that inference problem encountered remain tractable.
  \end{itemize}

% \textbf{There is no doubt that implementing some fo these features will be more invasive to the underlying simulator code than previously
% just hijacking the raw-random number draws, but, it would enable us to attack difficulties in the inference procedures that are
% currently not amenable to the current approach.}

\section{Sub-Programs within Simulators}

\begin{itemize}
  \item Talk about how within simulators we have nested-structures which are \emph{nested-programs}
  \item Define the notations used to define the processes within the simulator, in terms of the PPL framework
  \item Talk about the problems that such nested-programs cause to inference procedures
  \item Define the density of the simulator in terms of these sub-programs. 
\end{itemize}

Sub-programs, also called nested programs~\cite{rainforth2018nesting}, are an integral part of stochastic simulators.
Within stochastic simulators they are used to simulate densities that cannot be sampled from in any known, or efficient way.
Such nested-programs may be multi-tiered, in the sense that there exists several nested structures within a given sub-program. For example, given an outer sub-process $S_{out}$, that contains another sub-process $S_{in_{1}}$,
which itself contains another sub-program, $S_{in_{2}}$, that is $S_{in_2} \in S_{in_1} \in S_{out}$. 
Due to these recurrent dependency structures within simulators performing statistical inference becomes increasingly challenging, part of this
difficulty is due to the exploration of the domain of the density $R^{d}$. 
Because of the nested structures the inference strategy see a subset of the domain. 
This happens due to stochastic control flow within the sub-processes, which means only a subset of paths are available 
during the evaluation of the density. 
\textbf{Talk about why this causes trouble for the inference procedures etc}

Due to these complexities and the increase in computation time for both inference, \texttt{.backward()}, and the simulation, \texttt{.forward()},
we leverage the use of amortization schemes to, as we show in Section~\ref{sec:amortizesubp}, correctly utilise MCMC inference procedures and reduce
computation time.

Each nested process representing a different density, each simulating some statistical process.

% To construct the mathematical framework in this setting we employ the notation of~\cite{rainforth2017automating}, 
% where \emph{raw samples} are represented by $ f_{a_i}(x_i | \eta)$ and can be thought of as prior terms.

% \textbf{To finish - Linked back to figure 1}
% In order to correctly characterize this problem we first provide a useful diagram demonstrating 
% the density of the system~\ref{fig:systemoverview}. \textbf{Lots to do}

 \input{thesystem.tex}

% In order to correctly characterize this problem we first provide a useful diagram demonstrating 
% the density of the system.

%  \input{thesystem.tex}

\section{Amortized Sub-programs}

\label{sec:amortizesubp}
One common problem in traditional simulators, \emph{programs}, outside of the standard probabilistic
programming framework is that they rely heavily on rejection sampling 
to generate samples from the target density. 
Not only is rejection sampling 
computational inefficient, [especially if the criteria in which samples are generated
has been poorly calibrated] (may leave out this statement as it may make a reviewer
question why would amortization work any better.), but it creates complications for inference.
As the rejection sampling loops provide no explicit form of the target density,
we can not simply evaluate the density and leverage statistically and computationally
efficient algorithms such as MCMC sampling. This inhibits our ability to transform an arbitrary
stochastic simulator into a probabilistic programming framework~\cite{baydin2018efficient}. 
To this end, we derive both a mathematical and engineering based framework to facilitate 
the amortization of rejection sampling loops in arbitrary simulators, which provides a path
way to both leverage a larger, more efficient class of inference methods and convert arbitrary
stochastic simulators into probabilistic programs systems.

\subsection{Rejection Sampling}

Rejection samplers are heavily used when designing simulators [ add SHERPA etc], due to their ease of use and ability to model 
any arbitrary distribution on $\mathbb{R}^{D}$, with a sufficiently well designed rejection sampling scheme~\cite{MEYER20083408,martino2012improved,casella}

There are two different forms of rejection samplers found with in simulators and program code. 
In both instances we we have $X$, a random variable whose density is $f(x)$ that we cannot easily 
sample from, but can evaluate.
We then define a proposal using a random variable $Z$, whose density is $g(z)$  that we can efficiently
sample from.
Where the two methods diverge is in how the acceptance probability is defined and how the samples are generated.
In the standard statistical setting, see Algorithm~\ref{alg:rsstats}, we generate a sample $X$ by sampling $Z$ and accept $Z$ with probability
$p_{accept} = \frac{f(x)}{Mg(x)}$, which we can do until acceptance. 
Here $M$ is a constant that has a finite bound, $M \in (1, \infty)$ and the likelihood ratio $\frac{f(x)}{g(x)}$ is defined over the 
support of $X$. 
$M$ must also satisfy the following inequality $f(x) \leq M g(x)  \forall x$. 
This implies that the support of $Z$ must also contained the support of $X$, that is
$g(x) > 0$ whenever $f(x)>0$. 
We construct this acceptance probability by sampling from a uniform distribution defined on the interval 
$(0,1)$, which the proof ultimately relies on to ensure that a sample generated from the rejection sampler directly 
corresponds to a sample from the true, normalized, target density $f(x)$. 
But, provided we can calculate $P_{\text{accept}}(\mathcal{D} \leq \mathcal{C})$ for the given density, or mass, function
 $\mathcal{D}$ and condition $\mathcal{C}$, we can also satisfy the standard proof with a little more work.
 However, as we will demonstrate, rejection samplers 
 need not take this form and arise in many nested structures within probabilistic programs~\cite{rainforth2018nesting} and simulators~\cite{smith2008towards,bershteyn2018implementation,gleisberg2009event}.
 More generally speaking we can write a rejection sampler in the form of Algorithm~\ref{alg:rssim}
\begin{wrapfigure}{L}{0.40\textwidth}
 \begin{minipage}{0.40\textwidth}
  \null 
   \begin{algorithm}[H]
   \caption{ABC: Rejection Sampling}
   \label{alg:rsstats}
      \begin{algorithmic}
       \State {At iteration $i$ s.t $i \geq 1$}
       \State $X_{i} \sim g(X_{i})$ 
       \State $U_{i} \sim \mathcal{U}(0,1)$
       \If{$U_{i} \leq M \frac{f(X_i)}{g(X_{i})}$}
       \State{ accept $X_i \sim f$}
       \Else
       \State{$i \gets i + 1$, repeat}
       \EndIf
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}%
\end{wrapfigure}

\begin{wrapfigure}{L}{0.40\textwidth}
  \begin{minipage}{0.40\textwidth}
   \begin{algorithm}[H]
    \caption{Simulator: Rejection Sampling}
    \label{alg:rssim}
      \begin{algorithmic}
        \State {At iteration $i$ s.t $i \geq 1$}
        \State $X_{i} \sim p(x)$ 
        \State $\mathbb{I}(g(x) > 0)$
        \State $ Y \sim p(X=x | \mathbb{I}(g(x) > 0))$
        \State accept $X_i$ with $p(\mathbb{I}(g(x) > 0))$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{wrapfigure}

Thus, the probability of generating a sample $X = x$ from the true distribution can be defined, using Bayes rule, 
as: 
\begin{equation}
  p(X=x| \mathbb{I}) = \frac{p(\mathbb{I} | X=x)p(X=x)}{p(\mathbb{I})} 
\end{equation}
where $p(\mathbb{I} | X=x) = \{0,1\}$.

However, in many of the non-standard ABC forms of the rejection samplers, the acceptance 
probability is completely intractable, or is very difficult to calculate. 
Thus, as useful rejection sampling is for generating samples from densities that we 
would not be able to sample from, it has many short comings. 
These are of course well known and it is the reason for developing more sophisticated 
sampling schemes, such as importance sampling based methods.
However, in the setting of such complex simulators, written over many years and
in convoluted code bases, there is no easy way to re-implement such sampling 
schemes.
Hence, to leverage existing frameworks we can make use of the probabilistic 
programming execution (PPX) protocols~\cite{baydin2018efficient}. 
These protocols allow one to exact all raw sample calls $f_{a_j}$ from the
simulator. This then enables one to easily construct a prior for the full program
$P(\mathcal{T} = x_{1:n_{x}}| \lambda)$. 
In addition to this, once such rejection schemes, \emph{nested programs}, are located, we have a number of 
primitives that we can leverage to isolate calls to those addresses during training, to collect the samples generated during 
\texttt{.forward()} runs of the simulator, 
and during the inference procedure replace calls to those addresses $a_j$ with a
learnt proposal $\gamma_{a_j}(x_j ; \eta_{a_j}, \phi_j ) $.

% In particular, to simulate proposals for certain distributions,
% such as truncated normal distributions, it can be very difficult to
% hand-craft a proposal such that the acceptance rate.

Thus, the ability to learn and replace such rejection
sampling blocks with learnt surrogate functions, that are expressive enough to 
reciprocate the behavior of the rejection sampling block provides the user with three things. 
\begin{itemize}
  \item It reduces the computational complexity of the simulator.  
  \item As we now have a direct functional form of the density, we can correctly leverage more computationally and sample efficient inference schemes. 
  \item By only amortizing sub-programs and not amortizing the whole simulator, the larger structure of the simulator remains interpretable, in the sense that we can 
  still decode the decisions that it is making.  
\end{itemize}

\begin{figure}
  % \begin{minipage}[c]{0.67\textwidth}
  \includegraphics[width=0.6\textwidth, height=0.6\textheight,keepaspectratio]{amoritized_rejection.pdf}
  % \end{minipage}\hfill
  % \begin{minipage}[c]{0.3\textwidth}
    \caption{Using PyProb we can locate all regions of inefficient sampling 
    and replace those points in the code with learnt functions that represent
    the distributions of the rejection sampling statements. $Z_{i}$ represents the marginal and 
    by learning surrogate functions for each of the rejection sampling loops.
    In doing so we can learn approximate raw sample distributions $f(x_{i} | \eta)$ for each
    rejection sampling loop. This means that we can evaluate the function at each part in the simulator, enabling
    us to construct a target density that can be evaluated when using MCMC inference schemes such as RMH.}
    \label{fig:amortized_sampling}
  % \end{minipage}
\end{figure}

% We do this by replacing the rejection sampling blocks with 
% learnt functions from an amortized inference procedure Figure~\ref{fig:amortized_sampling}.
% This then enables us to replace rejection sampling loops with amortized functions that 
% are quick to evaluate, provide a functional form to the density and significantly reduce 
% memory over-heads, as compared to~\cite{baydin2018efficient}, due to not having to track the 
% thousands of raw sample calls $f_{a_i}$ generated in standard simulators.

\section{A Stochastic Simulator}

In this section, we present a simple stochastic simulator and povide 
examples of the rejection blocks found within standard stochastic
simulators~\cite{smith2008towards,bershteyn2018implementation, gleisberg2009event}. 
This simulator will form the basis for our experiemnts in Section~\ref{sec:experiments}.
Consider the following program, that is representative of a simple stochastic simulator
and two types of rejection sampler:

\noindent
\begin{tabular}{|p{3.6cm}|p{3.6cm}|p{3.6cm}|p{3.6cm}|}

\hline
Stochastic Simulator  &  R1 & R2 & R3 \\
\hline
\begin{lstlisting}[mathescape=true]
    def s():
      $z_{1} \sim \mathcal{N}(0,1)$
      $z_{2} \gets R_{1}(z_1, 1)$
      $z_{3} \sim \mathcal{U}(0,2)$
      $z_{4} \gets R_{2}(z_2,z_3, 1)$
      $z_{5} \sim \mathcal{U}(50,30^2)$
      $z_{6} \sim R_{3}(z_5, 1)$
    return $g(z_{1:6})$
\end{lstlisting}&
\begin{lstlisting}[mathescape=true]
def $R_1(z_1)$:
  $x \sim \mathcal{N}(z_1,1)$
  if $x > 0$:
    return $x$
  else:
    return $R_{1}(z_1)$
\end{lstlisting}&
\begin{lstlisting}[mathescape=true]
def $R_2(z_2, z_3)$:
  $y \gets 0$
  while $y < z_2$:
    $y \sim \mathcal{N}(z_3,1)$
  end
  return $y$
\end{lstlisting}&
\begin{lstlisting}[mathescape=true]
def $R_3(z_5)$:
  $f(x) = \mathcal{N}(30,10^2) +$
  $\mathcal{N}(80,20^2)$
  $g(x) = \mathcal{N}(50,30^2)$
  $u \sim \mathcal{U}(0,M*g(z_5))$
  if $u \leq f(z_5)$:
    return $z_5$
  else:
    return $R_3(z_5)$
\end{lstlisting}\\
\hline
$P(\mathcal{T}= x_{1:n_x} | \lambda)$ & $q_{1}(z_{2} | z_{1})$ & $q_{2}(z_{4} |z_{2},z_{3})$ & $q(z_5)$ \\
\hline
\end{tabular}

Here the rejection sampling functions take an additional argument $1$, or $0$,
to specify whether or not we want to collect samples to train, $1$, else replace with 
the learnt function, $0$. $P$ represents the program density and $q_{1}$ and $q_{2}$ represents 
the unknown densities that we wish to learn through an amortized approach.
% It is important to note that the expected running time of the rejection sampler is expected to geometric mean, we can show this as follows:
% \textbf{Bradley:  You only arrive at this, if the density is uniform} i.e R3
% \begin{align*}
% P( X= x) = \sum^{\infty}_{n=1} P(\text{reject $n-1$ times}, \text{draw} Y = x) \\
% \text{todo} 
% \end{align*}

If we have direct access to a rejection sampling loop then it is indeed 
possible to construct the density, up to a constant of proportionality. 
% That is:
% \begin{align*}
% z \sim g(\theta) \\
% do something \\
% \alpha = p_{accept}(z,\theta) \\
% \mathtt{if} Bern(\alpha) = 1: \\
% return z
% \end{align*}
 $x,z \propto f(\theta) p(accept)$.
However, in order to do construct this we need the input $z$, the sample from the proposal $x$ and $p(accept)$ (should use the notation above).
For any rejection sampling block, written in a generic code base, not in the form of R3, this is an incredibly hard 
task to do, because rejection sampling blocks within simulators are not necessarily written in a standard template. 
\textbf{Add examples in a supplement.  Need to demonstrate to YW that this is challenging to do, even if you know where 
the rejection sampling loops are. Hypothesis, RS loops are not written in a templated form}.
y leveraging the PPX protocols previously developed in \cite{baydin2018efficient,gram2019hijacking} we can however, 
extract all variables, other than $P(accept)$ in an automated way.
As without being able to easily extract $p(accept)$ this means that we are unable to 
correctly build the form of the density and thus are limited in our ability to leverage 
powerful inference methods such as MCMC and Variational methods.
Thus, we propose another approach, automating the amortization of rejection sampling loops of type R1, R2 and R3. 
This enables us to learn functions that we can easily sample from and via the amortization process construct the density of the entire
simulator.

The simulated densities for the rejection sampling block can be seen in Figure~\ref{fig:rs_simulated}.
\begin{figure}
  \label{fig:rs_simulated}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=0.6\textwidth, height=0.6\textheight, keepaspectratio]{rs_plots.png}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=0.6\textwidth, height=0.6\textheight, keepaspectratio]{rs3_plot.png}
  \end{minipage}
  \caption{ Simulated densities for rejection sampling block 1, \emph{blue} and rejection
  sampling block 2, \emph{orange} and rejection sampling block 3, \emph{red}.}
 \end{figure}

In order to replace such rejection sampling blocks with learnt functions that we 
can efficiently sample from, we must correctly specify our variational objective.

\subsection{Constructing the Objective}

The framework we present in this section is generally applicable to any stochastic simulator 
environment, where there exists any form of rejection sampling blocks.
To perform the inferences that we require, we construct a variational approximation of the rejection sampling 
blocks via an amortization procedure
and note that the variational approximation in the context of our setting is specified by three things:
\begin{enumerate}
  \itemsep0em 
  \item Makes use of existing parts of the outer program including all deterministic operations.
  \item Form of the $\gamma_{a_j}(x_j ; \eta_{a_j}, \psi_j) \forall a_j \in S_r$. 
  \item There exists variational parameters $\eta_{a_j} \forall a_j \in S_r$.
\end{enumerate}
where $a_j \in \{1, \ldots, L\}$ corresponds to the set of all addresses, $S_r$ denotes the set of addresses $a_j$ that correspond directly to the locations 
of the rejection sampling blocks, that is $a_j \cup S_r = \{1, \ldots, L\}$, $\gamma_{a_j}$ denotes the variational approximation at address $a_j$, 
for example, $\gamma_{a_j} = \mathcal{N}(x_j ; \mu_{a_j}(\psi_j; \eta_{a_j}), \Sigma_{a_j}(\psi_j ; \eta_{a_j}))$, with $\eta_{a_j}$ representing the variational
parameters and $\psi_j$ representing all variables that input into the given program.

We start by noting that the prior, defined in the context of the outer samples is defined as: 
\begin{equation}
  p(x) = 
  \begin{cases}
    \prod^{n_{x}}_{j=1} f_{a_j}(x_j ; \psi_j) \texttt{ if } \mathcal{B}(x) = 1 \\
    0
  \end{cases}
\end{equation}
next, we state for a given program $\mathcal{P}$ we denote the proposal for the program as:
\begin{equation}
  q(\mathbf{x}; \kappa) = 
  \begin{cases}
  \prod^{n_x}_{j=1, j\notin S_{r}} f_{a_j}(x_j ; \psi_j)(\prod^{n_x}_{j=1, j \in S_{r}} \gamma_{a_j}(x_j; \eta_{a_j}, \psi_{j})) \texttt{ if } \mathcal{B}(x_{1:n_{x}} = 1)\\
  0
  \end{cases}
\end{equation}
where $\kappa = \{\eta_{a_j} ; a_j \in S_{r}\}$ are the variational parameters.
As we now have the density of the proposal and the density of the program we can construct a variational objective as follows:

using the information projection we construct the variational objective as follows:
\begin{align*}
  J(\eta) = KL(p(x) || q(x; \kappa)) \\
   = \int p(x)\ln\left(\frac{p(x)}{q(x;\eta)}\right) dx \\
  %  \int p(x')dx' \gets 1 \\
  %  &= \mathbb{E}_{p(x)}[\ln(p(x))]  - \mathbb{E}_{p(x)}[\ln(q(x;\eta))]\\
  %  &= \ln(1) -  \mathbb{E}_{p(x)}[\ln(q(x;\eta))]\\ % I do not agree with this line
  %  &= -\mathbb{E}_{p(x)}[\ln(q(x;\eta))] \\
  %  &= -\mathbb{E}_{p(x)}[\sum^{n_x}_{j=1}\ln(q_{a_j}(x_j;\eta))] \\
   \text{Let } a_{j} \in \{1, \ldots, L\} , S_{R} \subset \{1, \ldots, L\} \\
   \kappa^{*} = \operatorname*{argmin}_{\kappa} \mathbb{E}_{X \sim p(x)}[-\log(q(X=x; \kappa))] \\
   = \operatorname*{argmax}_{\kappa} \mathbb{E}_{X \sim p(x)}[\sum^{n_x}_{j=1, j \in S_r} \log( \gamma_{a_j}(X_j = x_j ; \eta_{a_j}, \psi_j))] \\
   \forall r \in S_r, \eta^{*}_r = \operatorname*{argmax}_{r} \mathbb{E}_{x \sim p(x)}[\sum^{n_x}_{j=1}\mathbb{I}(a_j =r) \log(\gamma_r (X_j = x_j ; \eta_r, \psi_j))] \\
   \nabla_{\kappa} J(\kappa) = \nabla_{\kappa} \mathbb{E}_{X \sim p(x)}[\sum^{n_x}_{j=1}\mathbb{I}(a_j =r) \log(\gamma_r (X_j = x_j ; \eta_r, \psi_j))] \\
   = \mathbb{E}_{X \sim p(x)}[ \sum^{n_x}_{j=1, j \in S_r} \nabla_\kappa \log(\gamma_r (X_j = x_j ; \eta_r, \psi_j)]\\
   \approx \frac{1}{N} \sum^{N}_{n=1} \sum^{n_x}_{j=1, j\in S_r} \nabla_{\kappa}\log(\gamma_r (x^{n}_j ; \eta_r, \psi_j))
  %  = \cancelto{\text{no dependence on } \eta}{\mathbb{E}_{p(x)}[\sum^{n_x}_{j=1,a_j \notin S_{R}}\ln(p(x))]}-\mathbb{E}_{p(x)}[\sum^{n_x}_{j=1,a_j \in S_{R}}\ln(q_{a_j}(x_j;\eta))]\\
\end{align*}

Thus, to define the objective for each subproblem $r  \in S_{r}$ as:
\begin{equation}
  \label{eq:objective}
  \nabla_{\eta_r} J(\kappa) \approx \frac{1}{N} \sum^{N}_{n=1} \sum^{n_x}_{j=1} \mathbb{I}(r=a_j) \nabla_{\eta_r}\log(\gamma_r (x^{n}_j ; \eta_r, \psi_j))
\end{equation}

where $x^{n}_{1:n_x} \stackrel{iid}\sim p(x)$.

where during training we extract samples from each forward run and train each rejection sampler
separately.

\section{Experiments: Amortizing Stochastic Processes}
\label{sec:experiments}
\paragraph{TODO list}
\begin{itemize}
  \item Implement baselines
  \item Implement plotting scripts and evaluation scripts
  \item testing code and infrastructure
  \item literature review on any methods to learn rejection sampling methods, 
  and conditional density estimation. See this paper for kernel approximation 
  methods~\cite{erraqabi2016pliable}. In addition to the methods listed in the 
  rejection sampling subsection. 
\end{itemize}

\subsection{Ground truth}

Because in this simple case we have direct access to $\alpha$, the rejection sampler acceptance probability,
we can calculate the true density up to a constant of proportionality. 
\textbf{I will add this tomorrow}. I had no idea about this, until my meeting with 
YW today.

\subsection{Baseline}
For gif plots see the plots folder in code$\backslash$amortized\_rs$\backslash$plots

As a first approach, we train a regressor to learn a function to map
between the input / output pairs that are outputted from the rejection
sampling blocks with in the simulator.

For $R1$ we have one input $z^{r1} \in \mathbb{R}^{1 \times 1}$ and one output $y^{r1} \in \mathbb{R}^{1 \times 1}$.
For $R2$ we have two inputs $z^{r2} \in \mathbb{R}^{2 \times 1}$ and one output $y^{r2} \in \mathbb{R}^{1 \times 1}$.
For training purposes we generate $n_{b}$ batches of $n_{t}$ samples for training. 
This means that $z^{r1}_{batch}\mathbb{R}^{n_{b} \times 1}$. 
For the loss function we use a simple mean squared error loss \textbf{We can try different loss 
functions and optimizers} and use the
neural network, $\Gamma(z, y)$, to learn a $\hat{\mu}$ and $\hat{\sigma}^{2}$ to parameterize 
a normal proposal. For $R1$, this means $\Gamma_{r1}(z,y) \rightarrow \hat{\mu_{r1}},  \hat{\sigma_{r1}}^{2}$  
and so the proposal $q(z_{1} | z_{2}) = \mathcal{N}(\hat{\mu_{r1}},  \hat{\sigma_{r1}}^{2})$. 
Once learned we can then efficiently sample from the parameterized normal distribution, which
represents the density of the sub-program represented by $R1$. 
This is then replaced in the simulator, during inference time.

\paragraph{observations} 
Using a standard $L2$-loss function $L_{2} = (y_{pred} - y_{true})^{2}$
As the functions are relatively simple the loss functions it gets t
to a loss of around 0.6 and stops decreasing. Although, this seems low, I don't think that it is
actually good, because that would mean on average, for each data point there is a loss of $\pm$ 0.6
which per data point is quite bad. That may be for a number of reasons, using MSE loss, using a simple
model architecture. It really struggles to model the tails. It may require a much larger batch input size. 
. Increasing the batch size helped to correct for the peaked density, however the behavior is still odd.
The peaked behaviour still does not disappear, this, I believe is due to the loss function averaging 
over all points.

After my meeting with Tom, by using a loss function the problem of the network collapsing on a particular value was always going to 
happen as we should be optimizing our function via some objective, rather than looking at point-wise differences.

We had defined an objective above but it was not clear how this was supposed to be implemented in practice with the given notations.

We discussed at length how to correctly define the objective, I have now updated the objective with the one derived in the revised discussion,
along with additional notes on rejection sampling.

 \textbf{Add results from current experiments}

\subsection{2nd Baseline}
Instead of using a single Gaussian, use a proposal that is a mixture of Gaussians. 
That is, for $R1$,   $q(z_{1} | z_{2}) = \sum^{N}_{i=1}K_{i}\mathcal{N}(\hat{\mu^{i}_{r1}},  \hat{\sigma^{i}_{r1}}^{2})$
under the constraint that $\sum^{N}_{i=1}K_{i} = 1$. Maybe $N$ can be treated as a hyper-parameter. 
Or choose particular values.
\subsection{More complicated: Conditional Density estimation}

\subsection{3rd Baseline}

Use the variational objective constructed above.

\paragraph{Normalizing flows} One recent technique for density estimation, in the amortized inference setting,
is normalising flows combined with variation inference~\cite{rezende2015variational}.
Flow-based methods work by leveraging a simple property of differential geometry applied 
to random variables. 
That being the change of variables. 
Using flow based methods we can construct our rejection sampling block density, by learning a flow to map from
a simpler distribution to more complicated one.
Let $z \sim r(z)$ where $r(z)$ represents the posterior defined by the rejection
sampling loop, and define $\pi$ to be a smooth and invertible function.
When $\pi$ acts $z$ we transform the random variable from one domain, to another. 
Thus, if $z\prime = \pi(z)$ then we can construct the density of  
the probability density of the new random variable $z\prime$ under the flow initiated by $\pi$: 
\begin{equation}
r(z\prime) = r(\pi^{-1}(z\prime)) \left|\det \frac{\partial \pi^{-1}}{\partial z'}\right| = r(\pi^{-1}(z\prime)) \left|\det \frac{\partial \pi}{\partial z\prime}\right|^{-1} 
\end{equation}

Flows available are Real-NVP~\cite{dinh2016density} and Neural Spline flows~\cite{durkan2019neural}.
There are several others, for other techniques ask Gunes.

\begin{itemize}
  \item Do we need to know the base distribution that the RS block is sampling from? i.e there maybe be multiple $r(z)$ in each 
  RS block. 
  \item Is it right to say here that $z\prime$ is the output of the RS block, we do not know $\pi$ here, or is $r(z)$ the raw sample $f_{a_i}$ and $\pi$ the distribution specified by the rejection sampling loop? 
  \item We need to learn $\pi(.)$
\end{itemize}
\subsection{Using variants of VAEs}

Notes from models that define an explicit,  
tractable density are highly effective, 
because they permit the use of an optimization algorithm directly on the log-likelihood of the training data.
However,  the  family  of  models  that provide  a  tractable  density  is  limited,
with  different  families  having  different disadvantage.
The rejection sampling loops, if proposed correctly, provide a way of generating 
samples from some density, but do not provide a way of constructing the target density (\textbf{Bradley: Ask Tom, is this true?}). 
Within event and population-based simulators, the rejection sampling loops have been designed with care,
to ensure that they are representative to the events that they are simulating. So we can, with confidence,
assume that they are representative of some density. 
However, to utilise powerful inference engines, we must be able to construct the 
full density of the program.

\section{The difference between this paper and Inference compilation paper}

The main fundamental difference is that IC cares about the full global program, 
we only care about the local (sub / nested programs).

We are not interesting in calculating the weights $w^k$ since the method that we 
develop here is applicable to any inference scheme, in theory.

In the IC paper you are approximating $p(x|y)$ the \emph{the global} program, where 
as you can view this work as learning an approximation for the \emph{local}, sub-programs
$P^{in}(x|\nu)$ where the nuisance parameters in this paper are represented by $\psi$.

In the notation of the inference compilation paper we essentially learn $q_{a_{t},i_{t}}(x_t| x_{1:t-1}, y)$
where the indices are part of the set $S_r$.
 The difference here though, is that we only care about the input at that address, your proposals 
 are conditioning on everything up to that point of the execution trace $\mathcal{T}$
 in the simulator. 
 The training data represented by (11) in the IC paper is not the training data we use in this paper,
 we only care about the $x^{m}_t$ or $z$'s and $x$'s in this paper (notation is not entirely consistent through out this paper yet)
 But these are the parameters that are generated from raw sample calls within side the RS loop $f_{a_i \in S_r}(\cdot))$.

Thus the density of the full joint of the program in this setting is given as:
\begin{equation}
  p(x,y) = \prod_{t=1, t \notin S_r} f_{a_t}(x_t | \psi_t) \prod_{n=1} g_{b_n}(y_n | \phi) \prod_{h=1}\gamma_{a_j \in S_r}(\cdot)
\end{equation}
\section{Friday morning}
The subprograms within the simulator should be treated as nested probabilistic programs.

There are three ways we can do these things, rather than just an amortization

\subsection{Method 1: Learn the amortization of the sub-programs}
That is learn $\gamma_{a_{j}}(\cdot ; \eta_{a_{j}}, \cdot)$ for each program in the set $S_r$.

As per we are currently planning. Relies on generating exact samples, which 
we have access due to the rejection sampling loops.

In this setting we can approximate the output
\subsection{Method 2: Proper amortized inference of sub-programs}

Learn a $q_{a_{j}}(z_{1:j}; \eta_{a_j}, \phi_j)$ and uses this as a proposal for the 
inner inference. 
I.e. if we learn that $q$ we could develop a more efficient rejection sampler by replacing the proposal.

I.e we are calibrating the proposal to make it more amenable to accepting samples. 
We can do this by using self normalized importance sampling on the inner program.

The likelihood is just a $\{0,1\}$ likelihood, ie Bernoulli.

\subsection{Method 3: Learn marginal likelihood approximation}
With $ p(\cdot)$ representing the marginal likelihood. 
\begin{equation}
  p(x) = 
  \begin{cases}
  \prod^{n_x}_{j=1,{a_j}\in S_{R}} f_{a_j}(x_j ; \psi_j) \prod^{n_x}_{j=1,{a_j}\in S_{R}} \frac{P^{in}_{a_j}(x_j,\psi_j)}{P^{in}_{a_j}(\psi_j)} \\
  0
  \end{cases}
\end{equation}
The normalising constant inside (in) the rejection sampler is intractable.

The things inside of the rejection sampler are proportional to the following:
\begin{equation}
  P_{a_j}^{in}(x_j ; \phi_j) \propto
  \begin{cases}
  \prod^{n^{z}_{a_j}} f^{in}_{i}(z_{ij} | \theta_i) \prod_{k=1}^{n^{y}_{b_k}}g^{in}_{k}(y_i ; \phi_{i}) \texttt{ if } \mathcal{B}(z_{1:n^{z}_{a_j}}) = 1 \\
  0
  \end{cases}
\end{equation}

Thus, we need a regression from $\phi_j \rightarrow P^{in}{a_j}(\phi_j) $.
We have access to pairs $\{\phi_j, \hat{P}^{in}_{a_j}(\phi_j)\}$ for which 
the expectation $\mathbb{E}[\hat{P}^{in}_{a_j}( \phi_j)| \phi_j] = P^{in}_{a_j} (\phi_j)$ where
$\hat{P}$ is \textbf{Play back convo to find out}

It is really important to ensure that these pairs of samples are unbiased.

If you minimise an $\mathcal{L}_2$ norm for this given problem then the optimal proposal natural arises, see~\ref{fig:tomfig} 
The training regime would still need to be non-bias.

\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{tom_derivation.jpg}
  \caption{Tom's derivation proving that the $\mathcal{L}_2$ norm in this problem set-up enables us to learn a proposal that exactly
  matches the true distribution.  }
  \label{fig:tomfig}
\end{figure}
\section{ Questions for Tom}

\textbf{New questions to ask Tom}

\textbf{Old questions for Tom}
\begin{itemize}
  \item I don't understand how the objective function is comparing the generated samples from the density 
  estimator to the ground truth, at least in the way it is currently defined? Answered - we updated the objective defined and added more comments
  \item All RS references I look at state that we must sample from a uniform distribution and
  then check if our proposed value is less than that uniformly sampled point, yet our RS do not have this. Answered - Rejection samplers we have are valid rejection samplers, this just might not define the density we expect them to. ..?
  nor what is the acceptance probability within our rejection sampling statements? The proof for rejection sampling seems to rely on the fact that you are drawing from a uniform distribution.

  \item Whilst the variational posterior we choose as our proposal has a functional form  
  and can be sampled, the functional form of the rejection sampling loops is not explicitly known,
  which leads to problems when calculating the $KL$-divergence - reparametrize a with a function from a family of easy to sample distribution. 
  \item There are two approaches that we can take to calculating some divergence metric. 
  \item 1) Using something like a non-parametric divergence estimator that makes use of Renyi and $L_{2}$ divergences (see \cite{basseville2013divergence} for details) - No point, more difficult to implement
  \item 2) Use some form of kernel density to estimation to approximate the functional form of the rejection sampling 
  block, which can be used in place. From which we learn a variational approximation characterized by a trained network that we can sample from quickly. 
  \item An issue with approach 2) is that if I can approximate the density in functional form aren't there more efficient methods that we can use? - There are other methods out there and it would maybe make sense to compare some of those methods. However, it may still be difficult to sample. 
  \item Or, if assuming that the rejection sampling block can be arbitrary dimension, are we first learning a flow to represent that density and then learn a variational approximation of something that we can easily sample from. 
  \item But, this is very much the idea behind normalising flows, take a simple distribution that you can sample and then map it to the ``complicated'' density - Question not well defined. 
  \item Stein Variational inference may be quite nice to use here~\cite{feng2017learning}, but it does make explicit assumptions in requiring the densities to be differentiable. - Yes this method is good, but only works in the continuous domain. 
  \item  For an overview of all variational methods, including the use of different divergence metrics see here ~\cite{8588399} - Answered - Quite finicky to set up. 
\end{itemize}

\section{Additional TODOs}

\begin{itemize}
\item Before implementing such tools into the simulator, spend time looking for simpler stochastic simulators 
that can be hijacked but have target densities that can be constructed analytically. 
\item If we are not utilizing the additional information stored in the trace, which is only used during the inference compilation
training procedure, then we do not need it all, as for importance sampling we only require the weights after each \texttt{forward()}
update. 
\item If we cannot find any pre-existing simulators that are amenable to this approach then, we need to create our own, with
a rejection sampling loop, that can be exploited by the new machinery. This would enable us to debug, provide a simple example 
for the paper and understand the feasibility of the entire approach. 
\end{itemize}

\bibliography{refs}
\bibliographystyle{iclr2020_conference}

%%%%%%% Appendix %%%%%%%%
\appendix

\section{Code Examples of Stochastic Processes in Simulators }

The following examples are taken from SHERPA~\cite{gleisberg2009event}
a million-line plus code base used for the modelling of complex physics
simulators at CERN.

\small\begin{table}[]
  \begin{tabular}{|l|l|l|l|}
  \hline
  \textbf{Code} & \textbf{Base} & \textbf{Timed} & \textbf{Stochastic Control Flow} \\ \hline
  \begin{cpp}
  double Splitter_Base::SelectZ(const double & delta,
  const bool & lead)
   {double zmin(0.5*(1.-sqrt(1.-delta))),
   zmax(0.5*(1.+sqrt(1.-delta))), z;
    do {
      z = zmin+ran->Get()*sqrt(1.-delta);
        } 
    while (1.-2.*z*(1.-z) < ran->Get());
  return z;
   }
  \end{cpp}&Yes&No&Yes\\ 
  \hline
  \begin{cpp}
double Decay_Channel::GenerateMass(const double& max,
const double& width) 
const
{
 double mass=-1.0;
 double decaymin = MinimalMass();
 if(decaymin>max) mass=-1.0;
 else if (decaymin==0.0) {
   mass=m_flavours[0].RelBWMass(decaymin, max,
                p_ms->Mass(m_flavours[0]), width);
      }
 else {
   double s=sqr(p_ms->Mass(GetDecaying()));
   double mb(0.0), mc(0.0);
   for (int i=0; i<NOut(); ++i) {
     mc+=p_ms->Mass(GetDecayProduct(i));
     if(p_ms->Mass(GetDecayProduct(i))>mb)
       mb=p_ms->Mass(GetDecayProduct(i));
   }
   mc-=mb;
   double b=sqr(mb);
   double c=sqr(mc);
   double spmax=2.0*b+2.0*c+\\
   sqrt(sqr(b)+14.0*b*c+sqr(c));
   double wmax=MassWeight(s,spmax,b,c);
   double w=0.0;
   int trials(0);
   do {
     mass=m_flavours[0].RelBWMass(decaymin, max,
     p_ms->Mass(m_flavours[0]), width);
     double sp=sqr(mass);
     w=MassWeight(s,sp,b,c);
     ++trials;
   } while (w<ran->Get()*wmax && trials<1000);
 }
 return mass;
}
\end{cpp}
        &  No             & Yes                     & No  \\ \hline
                &               &                     &                                  \\ \hline
  \end{tabular}
  \end{table}
\section{The unconditional acceptance probability}

Given a collection of i.i.d random variables $\{X_{i}\}_{i = 1:n}$ with 
cumulative density function (CDF) $F_{X}$. Then by the Glivenko-Cantelli theorem~\cite{tucker1959generalization} 
we can define the empirical distribution function for $x \in \mathbb{R}$ as:
\begin{equation}
  F_{n}(x) = \frac{1}{n}\sum^{n}_{i=1} \mathbb{I}_{[X_{i}, \infty)}(x)
\end{equation}
And for any fixed $x \in \mathbb{R}$, the Strong law of large numbers ensures that 
$F_{n}(x) \rightarrow F_{X}(x)$ as $n \rightarrow \infty$ as 
\begin{equation}
\mathbb{E}[\mathbb{I}_{[X_{i},\infty)}(x)] = 
 P[\mathbb{I}_{X_{i},\infty)}(x) = 1] = P[X_{i}\leq x] = F_(X)(x)
\end{equation}
where $\mathbb{I}_{A}(x)$  is the indicator function for a set $A$.

\begin{align}
  \text{Prob}(U \leq \frac{p_{X}(z)}{Mg_{Z}(z)} = \mathbb{E}_()
\end{align}
\section{Law of total expectations}

If $X$ is a random variable whose expected value is $\mathbb{E}[X]$ and $Z$ is a random variable on the same probability space
then: \begin{equation}
  \label{eq:lawoftotalexp}
  \mathbb{E}[X] = \mathbb{E}_{p(z)}[\mathbb{E}_{p(x|z)}[X | Z]]
\end{equation}
That is, the expected value of $X$ given $Z$ is the same as the expectation of $X$.

\section{Limitations with different inference engines}

\subsection{Prior based sampling}
In prior based sampling we take our existing program and look directly at the product of all the raw sample calls, denoted $f_{a_i}(x_i | \eta)$ and 
construct a proposal that is is dependent on those. In this sampling scheme our proposal is defined as:
\begin{equation}
  q(x_{1:n_{x}}) = 
  \begin{cases}
    \prod_{i=1}^{n_x} f_{a_i}(x_i | \eta) \hspace{0.1cm} \texttt{if} \hspace{0.1cm} \mathcal{B}(x_{1:n_{x}})=1 \\
    0
  \end{cases}
\end{equation}
where $\mathcal{B}(\cdot)$ is satisfied by the simulator and $f_{a_i}(\cdot)$ represents a
raw random sample, $\eta$ is a subset of the variables in-scope at the point of sampling
and $n_{x}$ is the number of latent variables.
 This is true in all inference engines used in this hijacking framework.
However, this makes use of no conditioning and is not sufficient for performing inference in
complex, population-based simulations.

\subsection{Non-prior Based Sampling}
In non-prior based sampling we utilise conditioning too, which enables 
us to explore spaces more efficiently. 
In the current hijacking set-up, we
can only do conditioning at the end of each \texttt{forward()} run of the simulator. 
This does inhibit certain inference engines and makes aspects of performing inference 
within this setting more complicated. 
Nonetheless, there are several inference schemes 
that we can exploit and utilise. 
We shall now give an Introduction to each of these schemes.

\subsection{Importance Sampling}
In the generalised probabilistic programming setting an importance sampling scheme over the 
program raw samples and observations can be defined as follows:
\begin{equation}
  \label{eq:posterior}
  p(\mathcal{T}=x_{1:n_x}| \lambda)=
  \begin{cases}
  \prod_{i=1}^{n_x} f_{a_i}(x_i | \eta) \prod_{j=1}^{n_y} g_{b_j}(y_j | \phi) \hspace{0.1cm} \texttt{if} \mathcal{B}(x_{1:n_x}) = 1\\
  0
  \end{cases}
\end{equation}

where the observe statements are defined by $g_{b_j}(.)$, $\phi$ plays the same role as 
$\eta$ and $n_y$ represents the number of observations.
The expectation over the whole program can be defined as: 
\begin{equation}
\mathbb{E}_{p(\Omega | \lambda)}[h(\Omega)] = \mathbb{E}_{p(x_{1:n_x} | \lambda)}[h(P_r(x_{1:n_x}))]
\end{equation}
\textbf{Note: It is a little weird how we define this, as the the program output $\Omega$ should
also include the product of the weights}

where $r$ represents a run of the program, $P_{r}(\cdot)$ is the entirety of the program output,
$h(\cdot)$ is just a deterministic function.

As the importance sampler updates the weights according to the following formula:
\begin{equation}
  w(x_{1:n_x}) = \prod^{n_x}_{i=1}\frac{f_{a_i}(x_i | \eta_i)}{q_i(x_i |f_{a_i}, \eta_i)} \prod^{n_y}_{j=1} g_{b_j}(y_j | \phi_j)
\end{equation}

When the number of observations is fixed, the combined weight is equal to $w(x_{1:n_x}) = \prod^{n_y}_{j=1} w_j $. 
To generate the weight up to each observe we can define $K(j)$ to be a function that returns the least $i$ before hitting an observe $j$.
This enables us to write the $j$-th weight as follows:
\begin{equation}
  w_j(x1:K(j)) = g_{b_j}(y_j | \psi_j) \prod^{K(j)}_{K(j-1)+1} v_i
\end{equation}
where we define $v_j = \frac{f_{a_j}(x_j | \eta_j)}{q_j(x_j |f_{a_j}, \eta_j)}$

Taking these definitions we can construct a target density over the 
latents that is proportional to the products of \texttt{sample} and \texttt{observe}
statements \textbf{Note: Add update posterior here}.
Thus, given the proposal $q(x_{1:K(t)})$ we define the importance weight as:
\begin{multline}
  w(x_{1:t}) = \prod^{t}_{j=1} w_j (x_{1:K(j)}) = \\
   \frac{p_t(x_{1:K(t)})}{q(x_{1:K(t)})} =
 \begin{cases}
  \prod^{t}_{j=1}g_{b_j}(y_j | w_i)\prod^K(j)_{K(j-1)+1} v_i \hspace{0.1cm} \texttt{if} \hspace{0.1cm} B_t(x_1:K(t))=1 \\
  0
 \end{cases}
\end{multline}

One problem with importance sampling is in its ability to scale and condition upon multiple observations. |
Choosing a \emph{good} prior remains challenging.

\subsection{Sequential Monte Carlo}
Given multiple observations, we can ake the inference more computationally efficient 
by employing sequential monte carlo (SMC), although as we can only condition at the end 
of one full run of the simulation, computational speed increases are minimal.
SMC allows for conditioning on multiple observations, but the number of observations must
be fixed. \textbf{Note: If we could condition on observations within the simulator, we 
could get dramatic speed increases over standard importance sampling.We can actually do this 
in pyprob, but it requires knowing where to condition. That can be quite challenging 
for arbitrary code. }

\begin{algorithm}
    \caption{SMC for Population Based Simulators - well any general program - not new}
    \label{alg:euclid}
    \begin{algorithmic}[1] % The number tells where the line numbering should start
        \Procedure{SMC}{$T,M$,\textbf{x}, \textbf{w},\textbf{v}, \textbf{g}}
          \For{$t = 1:T$}
            \For{$m = 1:M$}
              \State $w^{m} = 1  $
              \For{$k = K(t-1)+1:K(t)$}
                \State $x^m_k \sim q_k(x^m_k | f_{a^{m}_{k}},\eta^{m}_{j})$
                \State $w^m_t \gets w^t_m v_j^m(x_k^m)$
                \EndFor
            \EndFor
          \State $w^m_t \gets w^m_t g_{b_{t}^m}(y_t^m | \psi^m_t)$
          \State $\{x^{1:m}_{1:t}\}$ $\gets$ \texttt{resample}( $\{x^{1:m}_{1:t}\}$,  $\{w^{1:m}_{1:t}\}$ )
        \EndFor
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

To implement SMC in the current set-up we round require a forking method and the

\subsection{Random-walk Metropolis Hastings}

Due to the current set-up performing Random-walk Metropolis Hastings (RMH) is not feasible 
as the rejection sampling blocks within the simulator mean that we cannot correctly perform RMH
as we do not have a functional form for all parts of the simulator to construct a target density
that we can evaluate. \textbf{Note:
Ask tom what:  one can sometimes achieve improvements by instead using the type of the distribution object 
associated with $x_m$ (what does this first part mean?) 
 to automatically construct a valid random walk proposal that allows for improved
 hill climbing behavior on the individual updates. }.

\textbf{TODO: Add some of the plots generated from this model and an accompanying analysis to them.}

\section{Stochastic Variational Inference}

Using the standard notations we define the joint as follows:
\begin{equation}
  p(x, z ; \theta) = p_{\theta}(x,z) = p_{\theta}(x| z)p_{\theta}(z)
\end{equation}

We assume that the distributions that create the \emph{model} $p_{\theta}(x, z)$ can be:
\begin{enumerate}
  \item We can sample from each $p_i$
  \item We can compute the pointwise log pdf of $p_i$
  \item $p_i$ is differentiable w.r.t the parameters $\theta$
  \end{enumerate}

\subsection{Learning the model}

Our objective will either want to maximize, or minimize, depending on what we want to discover.

\begin{equation}
  \theta_{\text{min} \backslash \text{max} } = \text{argmax$\backslash$min}_{\theta}\log p_\theta(x)
\end{equation}

where the log evidence $p_\theta(x)$ is given by:
\begin{equation}
  \log p_{\theta}(x) = \log \int dz p_\theta(x,z)
\end{equation}

This is a doubly hard problem as even if $\theta$ is fixed the integral over the set
of latent variables defined by $z$ is intractable. 
Even if we know how to calculate the log evidence for all $\theta$, maximising or minimising, 
the log evidence as a function of $\theta$ is in general a difficult non-convex problem.

In addition to finding the parameters $\theta$ we also require the posterior over
the latent variables $z$:
\begin{equation}
  p_{\theta_{\text{max$\backslash$min}}}(z|x) =  \frac{p_{\theta}(x,z)}{\int dz p_{\theta_{\text{max$\backslash$min}}}(x,z)} 
\end{equation}

To this in the variational inference setting we learn an approximation to $p_{\theta_{\text{max$\backslash$min}}}(z|x)$
denoted by $q_\phi(z)$ where $\phi$ are the variational parameters.

\section{The Objective}

In many of these settings the ELBO loss characterises the problem that we 
aim to solve.

\begin{equation}
  ELBO = \mathbb{E}_{q_{\phi}(z)} [ \log p_{\theta}(x,z) - \log q_{\phi}(z) ]
\end{equation}
and we have for all choices of $\theta$ and $\phi$ that:
\begin{equation}
  \log p_{\theta}(x) \geq ELBO 
\end{equation}
\section{An Example: Modelling Parasite Densities in Simulation}
% \begin{wrapfigure}{r}{0.4\textwidth}
  \begin{figure}[h!]
    \centering
    \label{fig:parasite}
    \includegraphics[width=0.4\textwidth]{parasite_model.pdf}
    \caption{The DAG of the Parasite model.}
  % \end{wrapfigure}
  \end{figure}
The parasite model by Smith el al.~\cite{smith2006relationship} is a simple stochastic 
model for determining the number of infections within a given population demographic.
The models hypothesized in ~\cite{smith2006relationship} were found, 
after extensive testing and development, to not characterise well the number 
of infections among different demographics, especially those under 15 years old. 
By using probabilistic programming we can immediately understand 
the expressiveness of the proposed model and can exact important insights for 
extending the proposed models, which increases ones ability to develop 
improved models to estimate parasite densities, given 
our observations of the seasonal entomological inoculation rate (EIR).
In addition to this we can greatly reduce the computational costs, whilst
improving model comprehension and expressivity.

\paragraph{Common features of the models}
The models used are all discrete time micro-simulations of malaria in humans, 
based on the published model used in[1], which we refer to as the base model.
The model for infection of the vector as a function of parasite densities[2] 
(and hence for the effect of vaccination on onward transmission), the case-management model[3], 
the model of vaccination by pre-erythrocytic vaccines[4], the models for clinical outcomes and mortality[5,6], 
and the basic design of the scenarios used for predicting impacts[3,7] all correspond to the previous descriptions,
 which are accompanied by detailed rationales for the formulations used.

For all models, the number of infections  introduced in individual $i$ in five-day time step $t$, is distributed as:
\begin{equation}
  h(i,t) \sim \textit{Pos}(\lambda(i,t))
\end{equation}
where the expected number $\lambda(i,t) = S(i,t)E_a(i,t)$ and $S(i,t)$ is the suseptability
(proportion of inoculations that result in infections) and and $E_{a}(i,t)$ is the expected number of entomological inoculations, adjusted for age and individual factors.

\paragraph{Base model}
Although a statistical model is not formally defined within~\cite{smith2006relationship}
we define the unknown \emph{disease dynamics} parameters with a set of latent variables$\{x_{i}\}^{N}$
and construct the model as follows:

In the base model $E_{a}(i,t) = E_{max}(t) \frac{A(a(i,t))}{A_{max}}$ where $E_{max}(t)$ refers to the usual measure of the EIR computed from human bait collections on adults $A(\cdot)$ and , the availability to mosquitoes,
 is assumed to be proportional to average body surface area and to depend only on age $a(i,t)$.
The function used to compute $A(a(i,t))$ increases with age up to age 20 years where it reaches a value of $A_{max}$.

To capture observed relationships between infection and measured entomological exposure, the susceptibility is then:
\begin{equation}
  S(i,t) = (S_{\infty} + \frac{1 - S_{\infty}}{1 + \frac{E_a(i,t)}{E^{*}}})(S_{imm} + \frac{1 - S_{imm}}{1 + (\frac{X_p(i,t)}{X_{p}^{*}})^{\gamma_p}})
\end{equation}
where $S_{imm}, X^{*}_{p} E^*, \gamma_p, S_\infty$ are latent variables
and $X_p(i,t) = \int^{t}_{-a(i,t)} E_{a}(i,\tau) d\tau$
Fitting this model to data of~\cite{beier1994plasmodium} using their method learns the values $S_{\infty} = 0.049$ and $E^* = 0.032$ 
inoculations/person-night. 
The model for each individual infection $j$ in host $i$ comprises a time series of parasite densities.
  In a naïve host the expectation of the logarithm of the density of the infection at time point $\tau$
in the course of the infection, $\ln(y_0(i,j,\tau))$, is taken from a statistical description of parasite densities
in malaria therapy patients~\cite{maire2006model}. 
In the presence of previous exposure and co-infection, the expected log density for each concurrent infection is: 
\begin{equation}
  E(\ln(y(i,j,\tau))) = D_{y}(i,t)D_{m}(i,t)D_{h}(i,t)\ln(y_0(i,j,\tau)) + \ln(\frac{D_x}{M(i,t)} + 1 - D_x))
\end{equation}
where $M(i,t)$ is the total multiplicity of infection and 
\begin{equation}
  D_{y}(i,t) = \frac{1}{1 + \frac{X_y(i,j,t)}{X^{*}_{y}}}
\end{equation}
where $X_y(i,j,t) = \sum^{t}_{t-a}Y(i,t) - \sum^{t}_{t_{0,j}} y(i,j,\tau)$ (note that a continuous time approximation to this
is given in the original publications~\cite{smith2006mathematical,maire2006model}
and hence measures the cumulative parasite load, where
\begin{equation}
  D_{y}(i,t) = \frac{1}{1 + \frac{X_{h}(i,t)}{X^{*}_h}}
\end{equation}
where $X_h(i,t) = \sum^{t}_{t-a}h(i,\tau) - 1$, the number of inoculations since birth, excluding
the one under consideration, which measures the inocula experienced by the host
up to the time point under consideration. 
\begin{equation}
  D_{m}(i,t) = 1 - \alpha_m \exp(-\frac{0.693a(i,t)}{a^{*}_m})
\end{equation}
which measures the effect of maternal immunity. 
$X^{*}_{y}$, $X^{*}_h$, $D_x$, $a^*_{m}$ and $\alpha_m$ are essentially 
latent variables to be learnt given the field data.

\paragraph{Mass action model of the force of infection}
The base model assumes that susceptibility is a sigmoidal function of the exposure, which conflicts
with the usual mass action assumption for infectious disease transmission.
However, if entomological exposure is over-dispersed, 
infection-exposure relationships similar to those observed in the field 
can be reconciled with mass action principles, thus providing explanation
for the observed non-linear relationship between infection and entomological exposure
and hence a more justifiable representation of the infection process than that used
in the base model~\cite{smith2008estimation}. 
We simulate this by including random variation in the availability of the human host to mosquitoes.
This variation comprises an individual-specific component which is captured by defining
for each simulated human a baseline availability, $A_b(i)$, which is sampled at birth using:
\begin{equation}
  \log(A_b(i)) \sim \mathcal{N}(-\frac{\sigma_b^{2}}{2}, \sigma^{2}_b)
\end{equation}

so that $A_b(i)$ is log-normally distributed with arithmetic mean 1. 
The age adhusted EIR at time $t$, adjusted to the individual is then:
\begin{equation}
  E_{b}(i,t) = E_{max}(t)A_b(t)\frac{A(a(i,t))}{A_{max}}
\end{equation}
Additional log normal variation is introduced at each time step in the simulation 
to make the expected number of entomological inoculations a function
both of the individual and of log-normal noise measured by the
variance parameter, $\sigma^2_n$ :
\begin{equation}
  \ln(E_{a}(i,t)) \sim Normal(\ln(E_{b}(i,t) - \frac{\sigma^2_n, \sigma^2_n}))
\end{equation}
The susceptibility in these models is independent of EIR, i.e.:
\begin{equation}
  S(i,t( = S_{\infty}(S_{imm} + \frac{1 - S_{imm}}{1 + (\frac{X_p(i,t)}{X_{p}^{*}})^{\gamma_p}})
\end{equation}

Three different parameterisations, XML files, of this model were considered in [8] (E0063,R0065 and R0068).
In each case, the overall variability in $E_{a}(i,t)$ was constrained to the same total,
estimated by fitting to the data of~\cite{beier1994plasmodium}.
A probabilistic programming approach to determining these paramteters would
have provided a way to leverage state of the art inference algorithms that 
enable conditioning on arbitrary models. 
The parameters learnt from the original approach taken by the Malaria community 
generated values of $\sigma_b^2 = 0.5, 0.3$ and $0.05$ for each model respectively, corresponding
to $\sigma^2_n = 0.187, 0.567$ and $0.704$. 
\textbf{TODO: To finish - Add prob prog example + plots}

\paragraph{Models of decay of natural immunity}
\textbf{To complete}

\end{document}